A High-Order Discontinuous Galerkin Solver with Dynamic Adaptive Mesh Refinement to Simulate Cloud Formation Processes

### 2019-10-31 writing/paper/paper.tex
Fix more stuff for PPAM 2019.

\documentclass[runningheads]{llncs}
\title{A High-Order Discontinuous Galerkin Solver with Dynamic Adaptive Mesh Refinement to Simulate Cloud Formation Processes }
\titlerunning{A High-Order DG Solver to Simulate Cloud Formation Processes}
\author{Lukas Krenz \and Leonhard Rannabauer \and Michael Bader}
\authorrunning{L.\ Krenz, L.\ Rannabauer, M.\ Bader}
\institute{Department of Informatics, Technical University of Munich\\
  \email{lukas.krenz@in.tum.de}, \email{rannabau@in.tum.de}, \email{bader@in.tum.de}
\vspace*{-0.2cm}
} % todo emails and stuff
\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage[american]{babel}
\usepackage[autostyle, english = american]{csquotes}
% \usepackage[%
%   backend=biber,
%   url=false,
%   doi=false,
%   % style=alphabetic,
%   % backref=true,
%   % hyperref=true,
%   maxnames=3,
%   % minnames=3,
%   % maxbibnames=99,
%   firstinits=true,
%   % uniquename=init
%   ]{biblatex}
% \addbibresource{../bibliography.bib}
\newcommand{\citeauthor}{???}

\usepackage{caption}
\usepackage{subcaption}
\usepackage{xparse} % for NewDocumentCommand
\usepackage{etoolbox} % for notblank (brackets only when argument)
\usepackage{xstring} % for \IfSubStr
\usepackage{xpatch}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools} % for \mathclap
\usepackage{nicefrac}
\usepackage{physics} % for derivatives
\usepackage{varioref}
\usepackage{nicefrac}
\usepackage{physics} % for derivatives
\usepackage[separate-uncertainty]{siunitx}
\usepackage{hyperref}
\usepackage[]{cleveref}
\newcommand{\creflastconjunction}{, and\nobreakspace} % use Oxford comma
\usepackage{todonotes}
\usepackage{multimedia}
\graphicspath{{../figures/}}
\let\boundary\undefined%
\crefformat{equation}{\eqA{}eq.~\eqB #2#1#3)}
% TODO: Fix for more than two equations!
\crefmultiformat{equation}{Eqs.~\eqMultiA#2#1#3\eqMultiB}%
{ and \eqMultiA#2#1#3\eqMultiB}{, \eqMultiA#2#1#3\eqMultiB}{ and~\eqMultiA#2#1#3\eqMultiB}
\newcommand{\eqA}{}
\newcommand{\eqB}{(}
\newcommand{\eqMultiA}{(}
\newcommand{\eqMultiB}{)}
\DeclareRobustCommand{\pcrefSingle}[1]{%
\begingroup%
  \renewcommand{\eqA}{(}\renewcommand{\eqB}{}%
\cref{#1}%
\endgroup%
}
\DeclareRobustCommand{\pcrefMulti}[1]{%
\begingroup%
    \renewcommand{\eqMultiA}{}\renewcommand{\eqMultiB}{}%
    (\cref{#1})%
\endgroup%
}
\DeclareRobustCommand{\pcref}[1]{%
\IfSubStr{#1}{,}{\pcrefMulti{#1}}{\pcrefSingle{#1}}%
}

@@

\begin{document}
\maketitle 
\begin{abstract}
  \todo{Corresp.\ author, ORCID next to authors}
  We present a high-order discontinuous Galerkin (\dg{}) solver of the compressible Navier-Stokes equations for cloud formation processes.
  The scheme exploits an underlying parallelized implementation of the \aderdg{} method with dynamic adaptive mesh refinement. 
  We improve our method by a \pde{}-independent general refinement criterion, based on the local total variation of the numerical solution.
  Our generic scheme shows competitive results for both classical \textsc{cfd} and stratified scenarios.
  While established methods use numerics tailored towards the specific simulation, our scheme works scenario independent.
  All together, our method can be seen as a perfect candidate for large scale cloud simulation runs on current and future super computers.\todo{Less bragging!}

\keywords{\aderdg{}  \and Navier-Stokes \and Adaptive Mesh Refinement}
\end{abstract}
\section{Introduction}
In this paper we address the resolution of basic cloud formation processes on modern super computer systems.
The simulation of cloud formations, as part of convective processes, is expected to play an important role in future numerical weather prediction~\cite{bauer2015quiet}.
This requires both suitable physical models and effective computational realizations. 
Here we focus on the simulation of simple benchmark scenarios~\cite{giraldo2008study}.
They contain relatively small scale effects which are well approximated with the compressible Navier-Stokes equations.
We use the \aderdg{} method of~\cite{dumbser2008unified}, which allows us to simulate the Navier-Stokes equations with a space-time-discretization of arbitrary high order.
In contrast to Runge-Kutta time integrators or semi-implicit methods, an increase of the order of \aderdg{} only results in larger computational kernels and does not affect the complexity of the scheme.
Additionally, \aderdg{} is a communication avoiding scheme and reduces the overhead on larger scale.
We see our scheme in the regime of already established methods for cloud simulations, as seen for example in~\cite{giraldo2008study,muller2010adaptive,muller2018strong}.

@@

\subsection{CDF Testing Scenarios}
We begin with a manufactured solution scenario which we can use for a convergence test.
We use the following constants of fluids for all scenarios in this section:
\begin{equation}
  \gamma = 1.4, \quad \Pr = 0.7, \quad c_v = 1.0.
\end{equation}
Our description of the manufactured solution follows~\cite{dumbser2010arbitrary}.
To construct this solution, we assume that
% \begin{align}\label{eq:manufactured-solution}
% \begin{split}
%   \pressure(\bm{x}, t) &= \pressure_0 \cos( \bm{k} \bm{x} - \omega t ) + \pressure_b, \\
%   \Qrho(\bm{x}, t) &= \Qrho_0 \sin (\bm{k} \bm{x} - \omega t) + \Qrho_b, \\
%   \Qv(\bm{x}, t) &= \bm{v_0} \sin(\bm{k} \bm{x} - \omega t),
% \end{split}
% \end{align}
\begin{align}\label{eq:manufactured-solution}
  \begin{split}
    \pressure(\bm{x}, t) &= \nicefrac{1}{10} \cos( \bm{k} \bm{x} - 2 \pi t ) + \nicefrac{1}{\gamma}, \\
    \Qrho(\bm{x}, t) &= \nicefrac{1}{2} \sin (\bm{k} \bm{x} - 2 \pi t) + 1, \\
    \Qv(\bm{x}, t) &= \bm{v_0} \sin(\bm{k} \bm{x} - 2 \pi t),
  \end{split}
  \end{align}
solves our \pde{}.
We use the constants $\bm{v_0} = \nicefrac{1}{4} \left( 1, 1 \right)^\intercal$, $\bm{k} = \nicefrac{\pi}{5} \left( 1, 1 \right)^\intercal$ and simulate a domain of size $\left[ 10 \times 10 \right]$ for $\SI{0.5}{\s}$.
The viscosity is set to $\mu = 0.1$.
% \begin{equation}
% \begin{alignedat}{6}
%   & \pressure_u &&= \nicefrac{1}{10}, \qquad& \pressure_b &= \nicefrac{1}{\gamma}, \\
%   & \Qrho_0 &&= \nicefrac{1}{2}, \qquad& \Qrho_b &= 1, \\
%   & \bm{v_0} &&= \nicefrac{1}{4},
%   \begin{pmatrix}
%     1 \\
%     1
%   \end{pmatrix},
%   \qquad& \bm{k} &= \nicefrac{\pi}{5}
%   \begin{pmatrix}
%     1\\
%     1
%   \end{pmatrix},\\
%   & \omega &&= 2 \pi.
%   \end{alignedat}
% \end{equation}
Note that \Cref{eq:manufactured-solution} does not solve the compressible Navier-Stokes equations \cref{eq:equation-set} directly.
It rather solves our equation set with an added source term which can be derived with a computer algebra system.
We ran this for a a combination of orders $1, \ldots 6$ and multiple grid sizes.
Note that by order we mean the polynomial order throughout the entire paper and not the theoretical convergence order.
For this scenario, we achieve high-order convergence (\cref{fig:convergence-test}) but notice some diminishing returns for large orders\todo{Maybe note that Dumbser doesnt get perfect convergence either.}.
\begin{figure}[tb]
  \centering
  \includegraphics{paper_convergence}
  \caption{\label{fig:convergence-test}Mesh size vs.\ error for various polynomial orders $P$.
    Dashed lines show the theoretical convergence order of $P+1$.}
\end{figure}

After we have established that the implementation of our numerical method converges, we are going to investigate three established testing scenarios from the field of computational fluid mechanics.
A simple scenario is the Taylor-Green vortex.
Assuming an incompressible fluid, it can be written as
\begin{align}
  \label{eq:taylor-green}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1,\\
  \Qv(\bm{x}, t) &= \exp(-2 \mu t)
  \begin{pmatrix}
    \phantom{-}\sin(x) \cos(y) \\
- \cos(x) \sin(y) 
    \end{pmatrix}, \\
  \pressure(\bm{x}, t) &= \exp(-4 \mu t) \, \nicefrac{1}{4} \left( \cos(2x) + \cos(2y) \right) + C.
  \end{split}
\end{align}
The constant $C = \nicefrac{100}{\gamma}$ governs the speed of sound and thus the Mach number $\text{Ma} = 0.1$~\cite{dumbser2016high}.
The viscosity is set to $\mu = 0.1$.

We simulate on the domain $[0,2\pi]^2$ and impose the analytical solution at the boundary.
A comparison at time $t = 10.0$ of the analytical solution for the pressure with our approximation (\cref{fig:taylor-green}) shows excellent agreement.
Note that we only show a qualitative analysis because this is not an exact solution for our equation set as we assume compressibility of the fluid.
This is nevertheless a valid comparison because for very low Mach numbers, both incompressible and compressible equations behave in a very similarly.
We used an \aderdg{}-scheme of order $5$ with a grid of $25^2$ cells.
%
\begin{figure}[tb]
  \centering
  \begin{subfigure}[t]{0.473\textwidth}
    \centering
    \includegraphics{paper_taylor_green_vel}
    \caption{\label{fig:taylor-green}Our result (markers) of the Taylor-Green vortex vs.\ the analytical solution (lines) \cref{eq:taylor-green}.
    The plot shows two velocity slices, the respective other coordinate is held constant at a value of $\pi$.}
  \end{subfigure}\qquad%
\begin{subfigure}[t]{.473\textwidth}
  \centering
    \includegraphics{paper_abc_flow_velocity}
    \captionof{figure}{\label{fig:abc-flow}Our approximation (markers) of the \textsc{abc}-flow vs.\ analytical solution (lines, \cref{eq:abc-flow}).
    All other axes are held constant at a value of 0. We show every 6th value.}
\end{subfigure}
    %\includegraphics{paper_abc_flow_velocity}
    %\caption{\label{fig:abc-flow}\textsc{abc}-flow. We show the velocity in $x$-direction for slices through the domain, where all but the mentioned coordinates are held constant at a value of zero.
    %The analytical solution is represented by lines, the numerical results by markers.}
  \caption{\label{fig:cdf-results}Two dimensional \textsc{cfd} scenarios}
\end{figure}

The Arnold-Beltrami-Childress (\textsc{abc}) flow is similar to the Taylor-Green vortex but is an analytical solution for the three-dimensional incompressible Navier-Stokes equations~\cite{tavelli2016staggered}.
It is defined in the domain \( \left[ -\pi, \pi \right]^3 \) as
\begin{align}
  \label{eq:abc-flow}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1,\\
  \Qv(\bm{x}, t) &= \phantom{-} \exp(-1\mu t)
  \begin{pmatrix}
    \sin(z) + \cos(y)\\
    \sin(x) + \cos(z)\\
    \sin(y) + \cos(x)
  \end{pmatrix}, \\
  \pressure(\bm{x}, t) &= -\exp(-2 \mu t) \, \left(\cos(x)\sin(y) + \sin(x)\cos(z) + \sin(z)\cos(y)\right)
  + C.
  \end{split}
\end{align}
The constant $C = \nicefrac{100}{\gamma}$ is chosen as before.
We use a viscosity of $\mu = 0.01$ and analytical boundary conditions.
Our results (\cref{fig:abc-flow}) show a good agreement between the analytical solution and our approximation with an \aderdg{}-scheme of order $3$ with a mesh consisting of $27^3$ cells at time $t = \SI{0.1}{\s}$.
Again, we do not perform a quantitative analysis as the \textsc{abc}-flow only solves our equation set approximately.

@@

\begin{figure}[tb]
  \centering
  \includegraphics[width=1.0\textwidth]{paper_two_bubbles}
  \caption{\label{fig:two-bubbles-ader}Left: Colliding Bubbles with \aderdg{}. Contour values for potential temperature perturbation are $-0.05, 0.05, 0.1, \ldots 0.45$.\\
  Right: Comparison of small scale structure between order 3 (top) and order 6 (bottom).}
\end{figure}
We have two bubbles, with constants
\begin{equation}
  \label{eq:bubbles-values}
\begin{alignedat}{6}
  & \text{warm:} \qquad && A = \SI{0.5}{\K}, \quad&& a = \SI{150}{\m}, \quad&& s = \SI{50}{\m}, \quad&& x_c = \SI{500}{\m,} \quad&& z_c = \SI{300}{\m},\\
  & \text{cold:} \qquad && A = \SI{-0.15}{\K}, \quad&& a = \SI{0}{\m}, \quad&& s = \SI{50}{\m}, \quad&& x_c = \SI{560}{\m}, \quad&& z_c = \SI{640}{\m}.
  \end{alignedat}
\end{equation}
Similar to~\cite{muller2010adaptive}, we use a constant viscosity of $\mu = 0.001$ to regularize the solution.
Note that we use a different implementation of viscosity than~\cite{muller2010adaptive}.
Hence, it is difficult to compare the parametrization directly.
We ran this scenario twice:
once without \amr{} and a mesh of size $\SI{1000/81}{\m} = \SI{12.35}{\m}$ and once with \amr{} with two adaptive refinement levels and parameters $\Trefine = 2.5$ and $\Tdelete = -0.5$.
For both settings we used polynomials of order 6.
We specialize the \amr{}-criterion~\pcref{eq:refinement-criterion} to our stratified flows by using the potential temperature.
This resulted in a mesh with cell-size lengths of approx.\ \SI{111.1}{\m}, \SI{37.04}{\m}, and \SI{12.34}{\m}.
The resulting mesh can be seen in \cref{fig:two-bubbles-ader}.
In addition, we ran a simulation with the same settings for a polynomial order of 3.
We can see~\pcref{fig:two-bubbles-ader} that the lower resolution leads to spurious wave and does not capture the behavior of the cloud.

We observe that the $L_2$ difference between the potential temperature of the \amr{} run, which uses 1953 cells, and the one of the fully refined run with 6561 cells, is only $1.87$.
The relative error is \num{2.56e-6}.
We further emphasize that our \amr{}-criterion accurately tracks the position of the edges of the cloud instead of only its position.
This is the main advantage of our gradient-based method in contrast to methods working directly with the value of the solution, as for example~\cite{muller2010adaptive}.
Overall, our result for this benchmark shows an excellent agreement to the previous solutions of~\cite{muller2010adaptive}.

In addition, we simulated the same scenario with our \muscl{} method, using $7^2$ patches with $90^2$ finite volume cells each.
As we use limiting, we do not need any viscosity.
The results of this method (\cref{fig:two-bubbles-fv}) also agree with the reference but contain fewer details.
Note that the numerical dissipativity of the finite volume scheme has a smoothing effect that is similar to the smoothing caused by viscosity.

@@

\subsection{Scalability}%
\newcommand{\mdofs}{M\text{DOF}/s}%
All two-dimensional scenarios presented in this paper can be run on a single workstation in less than two days.
Parallel scalability was thus not the primary goal of this paper.
Nevertheless, our implementation allows us to scale to small to medium scale setups using a combined \mpi{}~+~\tbb{} parallelization strategy, which works a follows:
We typically first choose a number of \mpi{} ranks that ensure an equal load balancing.
For two-dimensional simulations this happens for $1, 10, 83, \ldots$ ranks, as our underlying framework uses three-way splittings for each level and per dimension and an additional communication rank per level.
Then, we choose the number of nodes that we want to use and divide the cores into shared-memory sections.
These are then distributed among \tbb{} ranks.

We ran the two bubble scenario for a uniform grid with a mesh size of $729 \times 729$ for 20 timesteps and for multiple combinations of \mpi{} ranks and \tbb{} threads.
This simulation was performed on the SuperMUC-NG system using computational kernels that are optimized for its Skylake architecture\todo{Cite opt. kernels?}.
Using a single \mpi{} rank, we get roughly $4.9$ millions degrees of freedom updates ($\mdofs$) using two \tbb{} threads and $20.15 \mdofs$ using 24 threads (i.e.\ a half node).
For a full node with 48 threads, we get a performance of $12 \mdofs$.
When using 5 nodes with 10 \mpi{} ranks, we achieve $29.3 \mdofs$ for two threads and $137.3 \mdofs$ for 24 threads.

@@

\subsection*{Acknowledgments}
This work was funded by the European Union’s Horizon 2020 Research and Innovation Programme under grant agreements 
No~671698 (project ExaHyPE, \url{www.exahype.eu}) and 
No~823844 (ChEESE centre of excellence, \url{www.cheese-coe.eu}).
Computing resources were provided by the Leibniz Supercomputing Centre (project pr83no\todo{Mention other project?}).
Special thanks go to Dominic E.\ Charrier for his support with the implementation in the \exahypeengine{}.

>>>>

\documentclass[runningheads]{llncs}
\title{A High-Order Discontinuous Galerkin Solver with Dynamic Adaptive Mesh Refinement to Simulate Cloud Formation Processes }
\titlerunning{A High-Order DG Solver to Simulate Cloud Formation Processes}
\author{Lukas Krenz\orcidID{0000-0001-6378-0778} \and Leonhard Rannabauer \and Michael Bader}
\authorrunning{L.\ Krenz, L.\ Rannabauer, M.\ Bader}
\institute{Department of Informatics, Technical University of Munich\\
  \email{lukas.krenz@in.tum.de}, \email{rannabau@in.tum.de}, \email{bader@in.tum.de}
\vspace*{-0.2cm}
} % todo emails and stuff
\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage[american]{babel}
\usepackage[autostyle, english = american]{csquotes}
% \usepackage[%
%   backend=biber,
%   url=false,
%   doi=false,
%   % style=alphabetic,
%   % backref=true,
%   % hyperref=true,
%   maxnames=3,
%   % minnames=3,
%   % maxbibnames=99,
%   firstinits=true,
%   % uniquename=init
%   ]{biblatex}
% \addbibresource{../bibliography.bib}
\usepackage{cite}
\newcommand{\citeauthor}{???}

\usepackage{caption}
\usepackage{subcaption}
\usepackage{xparse} % for NewDocumentCommand
\usepackage{etoolbox} % for notblank (brackets only when argument)
\usepackage{xstring} % for \IfSubStr
\usepackage{xpatch}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools} % for \mathclap
\usepackage{nicefrac}
\usepackage{physics} % for derivatives
\usepackage{varioref}
\usepackage{nicefrac}
\usepackage{physics} % for derivatives
\usepackage[separate-uncertainty]{siunitx}
\usepackage{hyperref}
\usepackage[capitalise]{cleveref}
\newcommand{\creflastconjunction}{, and\nobreakspace} % use Oxford comma
\usepackage{todonotes}
\usepackage{multimedia}
\graphicspath{{../figures/}}
\let\boundary\undefined%
\crefformat{equation}{\eqA{}Eq.~\eqB #2#1#3)}
% TODO: Fix for more than two equations!
\crefmultiformat{equation}{Eqs.~\eqMultiA#2#1#3\eqMultiB}%
{ and \eqMultiA#2#1#3\eqMultiB}{, \eqMultiA#2#1#3\eqMultiB}{ and~\eqMultiA#2#1#3\eqMultiB}
\newcommand{\eqA}{}
\newcommand{\eqB}{(}
\newcommand{\eqMultiA}{(}
\newcommand{\eqMultiB}{)}
\DeclareRobustCommand{\pcrefSingle}[1]{%
\begingroup%
  \renewcommand{\eqA}{(}\renewcommand{\eqB}{}%
\cref{#1}%
\endgroup%
}
\DeclareRobustCommand{\pcrefMulti}[1]{%
\begingroup%
    \renewcommand{\eqMultiA}{}\renewcommand{\eqMultiB}{}%
    (\cref{#1})%
\endgroup%
}
\DeclareRobustCommand{\pcref}[1]{%
\IfSubStr{#1}{,}{\pcrefMulti{#1}}{\pcrefSingle{#1}}%
}

@@

\begin{document}
\maketitle 
\begin{abstract}
  % We present a high-order discontinuous Galerkin (\dg{}) solver of the compressible Navier-Stokes equations for cloud formation processes.
  % The scheme exploits an underlying parallelized implementation of the \aderdg{} method with dynamic adaptive mesh refinement. 
  % We improve our method by a \pde{}-independent general refinement criterion, based on the local total variation of the numerical solution.
  % Our generic scheme shows competitive results for both classical \textsc{cfd} and stratified scenarios.
  % While established methods use numerics tailored towards the specific simulation, our scheme works scenario independent.
  % All together, our method can be seen as a perfect candidate for large scale cloud simulation runs on current and future super computers.\todo{Less bragging!}
   We present a high-order discontinuous Galerkin (\dg{}) solver of the compressible Navier-Stokes equations for cloud formation processes.
  The scheme exploits an underlying parallelized implementation of the \aderdg{} method with dynamic adaptive mesh refinement. 
  We improve our method by a \pde{}-independent general refinement criterion, based on the local total variation of the numerical solution.
  While established methods use numerics tailored towards the specific simulation, our scheme works scenario independent.
  Our generic scheme shows competitive results for both classical \textsc{cfd} and stratified scenarios.
  We focus on two dimensional simulations of two bubble convection scenarios over a background atmosphere.
  The largest simulation here uses order 6 and 6561 cells which were reduced to 1953 cells by the refinement criterion.
 \todo{Abstract---ideas for last two sentences. Note that adding one sentence leads to more pages.}
% 
\keywords{\aderdg{}  \and Navier-Stokes \and Adaptive Mesh Refinement}
\end{abstract}
\section{Introduction}
In this paper we address the resolution of basic cloud formation processes on modern super computer systems.
The simulation of cloud formations, as part of convective processes, is expected to play an important role in future numerical weather prediction~\cite{bauer2015quiet}.
This requires both suitable physical models and effective computational realizations. 
Here we focus on the simulation of simple benchmark scenarios~\cite{giraldo2008study}.
They contain relatively small scale effects which are well approximated with the compressible Navier-Stokes equations.
We use the \aderdg{} method of~\cite{dumbser2008unified}, which allows us to simulate the Navier-Stokes equations with a space-time-discretization of arbitrary high order.
In contrast to Runge-Kutta time integrators or semi-implicit methods, an increase of the order of \aderdg{} only results in larger computational kernels and does not affect the complexity of the scheme.
Additionally, \aderdg{} is a communication avoiding scheme and reduces the overhead on larger scale.
We see our scheme in the regime of already established methods for cloud simulations, as seen for example in~\cite{giraldo2008study,muller2010adaptive,muller2018strong}.

@@

\subsection{CDF Testing Scenarios}
We begin with a manufactured solution scenario which we can use for a convergence test.
We use the following constants of fluids for all scenarios in this section:
\begin{equation}
  \gamma = 1.4, \quad \Pr = 0.7, \quad c_v = 1.0.
\end{equation}
Our description of the manufactured solution follows~\cite{dumbser2010arbitrary}.
To construct this solution, we assume that
% \begin{align}\label{eq:manufactured-solution}
% \begin{split}
%   \pressure(\bm{x}, t) &= \pressure_0 \cos( \bm{k} \bm{x} - \omega t ) + \pressure_b, \\
%   \Qrho(\bm{x}, t) &= \Qrho_0 \sin (\bm{k} \bm{x} - \omega t) + \Qrho_b, \\
%   \Qv(\bm{x}, t) &= \bm{v_0} \sin(\bm{k} \bm{x} - \omega t),
% \end{split}
% \end{align}
\begin{align}\label{eq:manufactured-solution}
  \begin{split}
    \pressure(\bm{x}, t) &= \nicefrac{1}{10} \cos( \bm{k} \bm{x} - 2 \pi t ) + \nicefrac{1}{\gamma}, \\
    \Qrho(\bm{x}, t) &= \nicefrac{1}{2} \sin (\bm{k} \bm{x} - 2 \pi t) + 1, \\
    \Qv(\bm{x}, t) &= \bm{v_0} \sin(\bm{k} \bm{x} - 2 \pi t),
  \end{split}
  \end{align}
solves our \pde{}.
We use the constants $\bm{v_0} = \nicefrac{1}{4} \left( 1, 1 \right)^\intercal$, $\bm{k} = \nicefrac{\pi}{5} \left( 1, 1 \right)^\intercal$ and simulate a domain of size $\left[ 10 \times 10 \right]$ for $\SI{0.5}{\s}$.
The viscosity is set to $\mu = 0.1$.
% \begin{equation}
% \begin{alignedat}{6}
%   & \pressure_u &&= \nicefrac{1}{10}, \qquad& \pressure_b &= \nicefrac{1}{\gamma}, \\
%   & \Qrho_0 &&= \nicefrac{1}{2}, \qquad& \Qrho_b &= 1, \\
%   & \bm{v_0} &&= \nicefrac{1}{4},
%   \begin{pmatrix}
%     1 \\
%     1
%   \end{pmatrix},
%   \qquad& \bm{k} &= \nicefrac{\pi}{5}
%   \begin{pmatrix}
%     1\\
%     1
%   \end{pmatrix},\\
%   & \omega &&= 2 \pi.
%   \end{alignedat}
% \end{equation}
Note that \Cref{eq:manufactured-solution} does not solve the compressible Navier-Stokes equations \cref{eq:equation-set} directly.
It rather solves our equation set with an added source term which can be derived with a computer algebra system.
We ran this for a a combination of orders $1, \ldots 6$ and multiple grid sizes.
Note that by order we mean the polynomial order throughout the entire paper and not the theoretical convergence order.
For this scenario, we achieve high-order convergence (\cref{fig:convergence-test}) but notice some diminishing returns for large orders\todo{Here a sentence regarding convergence would be nice---Dumbser got better results}.
\begin{figure}[tb]
  \centering
  \includegraphics{paper_convergence}
  \caption{\label{fig:convergence-test}Mesh size vs.\ error for various polynomial orders $P$.
    Dashed lines show the theoretical convergence order of $P+1$.}
\end{figure}

After we have established that the implementation of our numerical method converges, we are going to investigate three established testing scenarios from the field of computational fluid mechanics.
A simple scenario is the Taylor-Green vortex.
Assuming an \textit{incompressible} fluid, it can be written as
\begin{align}
  \label{eq:taylor-green}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1,\\
  \Qv(\bm{x}, t) &= \exp(-2 \mu t)
  \begin{pmatrix}
    \phantom{-}\sin(x) \cos(y) \\
- \cos(x) \sin(y) 
    \end{pmatrix}, \\
  \pressure(\bm{x}, t) &= \exp(-4 \mu t) \, \nicefrac{1}{4} \left( \cos(2x) + \cos(2y) \right) + C.
  \end{split}
\end{align}
The constant $C = \nicefrac{100}{\gamma}$ governs the speed of sound and thus the Mach number $\text{Ma} = 0.1$~\cite{dumbser2016high}.
The viscosity is set to $\mu = 0.1$.

We simulate on the domain $[0,2\pi]^2$ and impose the analytical solution at the boundary.
A comparison at time $t = 10.0$ of the analytical solution for the pressure with our approximation (\cref{fig:taylor-green}) shows excellent agreement.
Note that we only show a qualitative analysis because this is not an exact solution for our equation set as we assume compressibility of the fluid.
This is nevertheless a valid comparison because for very low Mach numbers, both incompressible and compressible equations behave in a very similarly.
We used an \aderdg{}-scheme of order $5$ with a grid of $25^2$ cells.
%
\begin{figure}[tb]
  \centering
  \begin{subfigure}[t]{0.473\textwidth}
    \centering
    \includegraphics{paper_taylor_green_vel}
    \caption{\label{fig:taylor-green}Our result (markers) of the Taylor-Green vortex vs.\ the analytical solution (lines) \cref{eq:taylor-green}.
    The plot shows two velocity slices, the respective other coordinate is held constant at a value of $\pi$.}
  \end{subfigure}\qquad%
\begin{subfigure}[t]{.473\textwidth}
  \centering
    \includegraphics{paper_abc_flow_velocity}
    \captionof{figure}{\label{fig:abc-flow}Our approximation (markers) of the \textsc{abc}-flow vs.\ analytical solution (lines, \cref{eq:abc-flow}).
    All other axes are held constant at a value of 0. We show every 6th value.}
\end{subfigure}
    %\includegraphics{paper_abc_flow_velocity}
    %\caption{\label{fig:abc-flow}\textsc{abc}-flow. We show the velocity in $x$-direction for slices through the domain, where all but the mentioned coordinates are held constant at a value of zero.
    %The analytical solution is represented by lines, the numerical results by markers.}
  \caption{\label{fig:cdf-results}Two-dimensional \textsc{cfd} scenarios}
\end{figure}

The Arnold-Beltrami-Childress (\textsc{abc}) flow is similar to the Taylor-Green vortex but is an analytical solution for the three-dimensional \textit{incompressible} Navier-Stokes equations~\cite{tavelli2016staggered}.
It is defined in the domain \( \left[ -\pi, \pi \right]^3 \) as
\begin{align}
  \label{eq:abc-flow}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1,\\
  \Qv(\bm{x}, t) &= \phantom{-} \exp(-1\mu t)
  \begin{pmatrix}
    \sin(z) + \cos(y)\\
    \sin(x) + \cos(z)\\
    \sin(y) + \cos(x)
  \end{pmatrix}, \\
  \pressure(\bm{x}, t) &= -\exp(-2 \mu t) \, \left(\cos(x)\sin(y) + \sin(x)\cos(z) + \sin(z)\cos(y)\right)
  + C.
  \end{split}
\end{align}
The constant $C = \nicefrac{100}{\gamma}$ is chosen as before.
We use a viscosity of $\mu = 0.01$ and analytical boundary conditions.
Our results (\cref{fig:abc-flow}) show a good agreement between the analytical solution and our approximation with an \aderdg{}-scheme of order $3$ with a mesh consisting of $27^3$ cells at time $t = \SI{0.1}{\s}$.
Again, we do not perform a quantitative analysis as the \textsc{abc}-flow only solves our equation set approximately.

@@

\begin{figure}[tb]
  \centering
  \includegraphics[width=1.0\textwidth]{paper_two_bubbles}
  \caption{\label{fig:two-bubbles-ader}Left: Colliding Bubbles with \aderdg{}. Contour values for potential temperature perturbation are $-0.05, 0.05, 0.1, \ldots 0.45$.\\
  Right: Comparison of small scale structure between order 3 (top) and order 6 (bottom).}
\end{figure}
We have two bubbles, with constants
\begin{equation}
  \label{eq:bubbles-values}
\begin{alignedat}{6}
  & \text{warm:} \qquad && A = \SI{0.5}{\K}, \quad&& a = \SI{150}{\m}, \quad&& s = \SI{50}{\m}, \quad&& x_c = \SI{500}{\m,} \quad&& z_c = \SI{300}{\m},\\
  & \text{cold:} \qquad && A = \SI{-0.15}{\K}, \quad&& a = \SI{0}{\m}, \quad&& s = \SI{50}{\m}, \quad&& x_c = \SI{560}{\m}, \quad&& z_c = \SI{640}{\m}.
  \end{alignedat}
\end{equation}
Similar to~\cite{muller2010adaptive}, we use a constant viscosity of $\mu = 0.001$ to regularize the solution.
Note that we use a different implementation of viscosity than~\cite{muller2010adaptive}.
Hence, it is difficult to compare the parametrization directly.
We ran this scenario twice:
once without \amr{} and a mesh of size $\SI{1000/81}{\m} = \SI{12.35}{\m}$ and once with \amr{} with two adaptive refinement levels and parameters $\Trefine = 2.5$ and $\Tdelete = -0.5$.
For both settings we used polynomials of order 6.
We specialize the \amr{}-criterion~\pcref{eq:refinement-criterion} to our stratified flows by using the potential temperature.
This resulted in a mesh with cell-size lengths of approx.\ \SI{111.1}{\m}, \SI{37.04}{\m}, and \SI{12.34}{\m}.
The resulting mesh can be seen in \cref{fig:two-bubbles-ader}.

We observe that the $L_2$ difference between the potential temperature of the \amr{} run, which uses 1953 cells, and the one of the fully refined run with 6561 cells, is only $1.87$.
The relative error is \num{2.6e-6}.
We further emphasize that our \amr{}-criterion accurately tracks the position of the edges of the cloud instead of only its position.
This is the main advantage of our gradient-based method in contrast to methods working directly with the value of the solution, as for example~\cite{muller2010adaptive}.
Overall, our result for this benchmark shows an excellent agreement to the previous solutions of~\cite{muller2010adaptive}.
In addition, we ran a simulation with the same settings for a polynomial order of 3.
The lower resolution leads to spurious waves~(\cref{fig:two-bubbles-ader}) and does not capture the behavior of the cloud.

Furthermore, we simulated the same scenario with our \muscl{} method, using $7^2$ patches with $90^2$ finite volume cells each.
As we use limiting, we do not need any viscosity.
The results of this method (\cref{fig:two-bubbles-fv}) also agree with the reference but contain fewer details.
Note that the numerical dissipativity of the finite volume scheme has a smoothing effect that is similar to the smoothing caused by viscosity.

@@

\subsection{Scalability}%
\newcommand{\mdofs}{\text{MDOF/s}}%
All two-dimensional scenarios presented in this paper can be run on a single workstation in less than two days.
Parallel scalability was thus not the primary goal of this paper.
Nevertheless, our implementation allows us to scale to small to medium scale setups using a combined \mpi{}~+~Thread building blocks(\tbb{}) parallelization strategy, which works a follows:
We typically first choose a number of \mpi{} ranks that ensure an equal load balancing.
\exahype{} achieves best scalability for $1, 10, 83, \ldots$ ranks, as our underlying framework uses three-way splittings for each level and per dimension and an additional communication rank per level.
For the desired number of compute nodes, we then determine the number of \tbb{} threads per rank to match the number of total available cores.

We ran the two bubble scenario for a uniform grid with a mesh size of $729 \times 729$ with order 6, resulting in roughly 104 million degrees of freedom (\textsc{dof}), for 20 timesteps and for multiple combinations of \mpi{} ranks and \tbb{} threads\todo{Check number of dofs? Should be $(6+1)^2 * 729^2$, right?}.
This simulation was performed on the SuperMUC-NG system using computational kernels that are optimized for its Skylake architecture.
Using a single \mpi{} rank, we get roughly $4.9$ millions \textsc{dof} updates ($\mdofs$) using two \tbb{} threads and $20.15 \mdofs$ using 24 threads (i.e.\ a half node).
For a full node with 48 threads, we get a performance of $12 \mdofs$.
When using 5 nodes with 10 \mpi{} ranks, we achieve $29.3 \mdofs$ for two threads and $137.3 \mdofs$ for 24 threads.

@@

\subsection*{Acknowledgments}
This work was funded by the European Union’s Horizon 2020 Research and Innovation Programme under grant agreements 
No~671698 (project ExaHyPE, \url{www.exahype.eu}) and 
No~823844 (ChEESE centre of excellence, \url{www.cheese-coe.eu}).
Computing resources were provided by the Leibniz Supercomputing Centre (project pr83no).
Special thanks go to Dominic E.\ Charrier for his support with the implementation in the \exahypeengine{}.
